---
title: "DB create and load"
output: html_document
---

## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dbplyr)
```

Make sure database file exists

```{bash}
touch ../data/gs_intensities.sqlite
```

Load the database and connection info
```{r}
library(DBI)
library(odbc)
library(config)
dsn <- get("gsint")
con <- dbConnect(odbc::odbc(),driver = dsn$driver, database = dsn$database , timeout = 10)
```

## Setup the database

The main design of the database is 3 tables: the main data table, 

### Create tables


Create the table for the markers
```{sql, connection = con}
CREATE TABLE IF NOT EXISTS "marker_info" (
  "markerid" INTEGER NOT NULL,
  "name" TEXT NOT NULL,
  "address" INTEGER,
  "chr" INTEGER NOT NULL,
  "position" INTEGER NOT NULL,
  primary key (chr, position, name)
);

```

create the table to store the info about the samples
```{sql, connection = con}
CREATE TABLE IF NOT EXISTS "batch"(
  "sampleid" CHAR NOT NULL,
  "batchid" INTEGER NOT NULL,
  "sex" CHAR,
  primary key (sampleid, batchid)
)
;
```


create the main data table
```{sql, connection = con}
CREATE TABLE IF NOT EXISTS "intensities"(
  "markerid" INTEGER NOT NULL,
  "sampleid" CHAR NOT NULL,
  "gtype" CHAR,
  "x_raw" REAL,
  "x" REAL,
  "y_raw" REAL,
  "y" REAL,
  FOREIGN KEY ("markerid") REFERENCES marker_info("markerid"),
  FOREIGN KEY ("sampleid") REFERENCES batch("sampleid")
);
```


## Data load

Load in the test data to R
```{r}
gs_tab <- read_delim(file = "~/Git_repos/gsQCplot/data/gs_export_table.txt",delim = '\t', n_max = 100) %>% set_names(., nm = str_replace(colnames(.), " ", "_")) %>% select(Index, Name, Address, Chr, Position, GenTrain_Score, starts_with("Frac_"), ends_with("GType"), ends_with(".X_Raw"), ends_with(".X"), ends_with(".Y_Raw"), ends_with("Y"))
```

```{r}
marker_info <- gs_tab %>% select(markerid = Index,name = Name, address = Address, chr = Chr, position = Position) %>% mutate_if(is.numeric, as.integer)
```

```{r}
db_insert_into(con, "marker_info", marker_info)
```

check markers were inserted
```{sql, connection = con}
SELECT * from marker_info limit 5;
```

make the samples into long format for the db

```{r}
suffixes <- c("GType", "X_Raw","X","Y_Raw","Y")
names(suffixes) <- suffixes
```


```{r}
gs_long <- purrr::map(suffixes, ~ gs_tab %>% select(Index, ends_with(paste0(".",.x))) %>% gather("sample", !!.x , ends_with(paste0(".", .x))) %>% mutate(sample = str_remove(sample, paste0("\\.",.x ))) )

```


join the results back together
```{r}
gs_long_combined <- purrr::reduce(gs_long, left_join, by = c("Index", "sample"))

```

populate the batch table with the sample info
```{r}
batch <- gs_long_combined %>% select(sampleid = sample) %>% mutate(batchid = 1, sex = NA) %>% distinct()
db_insert_into(con, "batch", batch)
```



```{r}
db_insert_into(con, "intensities", gs_long_combined %>% select(markerid = Index, sampleid = sample, gtype = GType, x_raw = X_Raw, x = X, y_raw = Y_Raw, y = Y))
```

```{sql, connection = con}
CREATE VIEW combined AS SELECT intensities.*, marker_info.name, marker_info.chr, marker_info.position FROM intensities LEFT JOIN marker_info on intensities.markerid = marker_info.markerid
```

